### 2013-07-03 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
adding pure domain decomposition write up

%%---------------------------------------------------------------------------%%
\section{Parallel MCSA\ }
\label{sec:parallel_mcsa}
With the parallel adjoint Neumann-Ulam solver implementation described
aboive, the parallel implementation of an MCSA iteration is
trivial. Recall the MCSA iteration procedure presented again here for
clarity.
\begin{subequations}
  \begin{gather}
    \label{eq:mcsa_r1}
    \ve{x}^{k+1/2} = \ve{x}^k + \ve{r}^k\:,\\
    \label{eq:mcsa_r2}
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\
    \label{eq:mcsa_r3}
    \ve{A}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:,\\
    \label{eq:mcsa_r4}
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}\:,\\
    \label{eq:mcsa_r5}
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:,
  \end{gather}
\end{subequations}
In \S\ref{sec:parallel_krylov_methods} we discuss parallel matrix and
vector operations as utilized in conventional Krylov methods. We
utilize these here for the parallel MCSA implementation for the
computations required in Eqs~(\ref{eq:mcsa_r1}), (\ref{eq:mcsa_r2}),
(\ref{eq:mcsa_r4}), and (\ref{eq:mcsa_r5}). In the first step, a
parallel matrix-vector multiply is used to apply the split operator to
the previous iterate's solution. A parallel vector update is then
performed with the source vector to arrive at the initial iteration
guess. In the next step, the residual is computed by the same
operations where now the operator is applied to the solution guess
with a parallel matrix-vector multiply and then a parallel vector
update with the source vector is performed. Once the correction is
computed with a parallel adjoint Neumann-Ulam solve, this correction
is applied to the guess with a parallel vector update to get the new
iteration solution. Additionally, as given by
Eq~(\ref{eq:mcsa_stopping_criteria}), 2 parallel vector reductions
will be required to check the stopping criteria: one initially to
compute the infinity norm of the source vector, and another at every
iteration to compute the infinity norm of the residual vector.

To parallelize Eq~(\ref{eq:mcsa_r3}), we apply the MSOD Neumann-Ulam
algorithm presented in the previous section. The problem is
potentially replicated with overlap and the Monte Carlo procedure
returns a correction vector, $\delta \ve{x}$, with the same parallel
decomposition as the input linear problem. As was the case for
generating the replications required for a multiple set
implementation, the explict matrix and vector operations required for
the rest of an MCSA iteration occur only on the subset of processors
containing the primary set. Only the Monte Carlo data structures exist
on the rest of the processors in the problem as replication of the
other operations does not generate any more information that could be
used to accelerate the time to solution.

@@

Of primary interest to the nuclear engineer is the \textit{strong
  scaling} performance of an algorithm
\citep{siegel_analysis_2012}. In this type of scaling study, the
global problem size is fixed and the local problem size changes with
the number of processes in the calculation. In this case, as $P$
increases, $N$ decreases, thus reducing performance by having a larger
ratio of communication to computational work. This type of scaling is
important because typically the global problem size is fixed due to
the problem of interest. Consider the desire for a high fidelity
neutronics simulation of the entire core of a nuclear reactor. In this
case, the global problem is known from the geometry definition
including the number of fuel assemblies and the geometry of those
assemblies all the way down to the resolution of the space, angle, and
energy discretizations required in order to capture physical phenomena
of interest. Because the global problem size is known a priori, strong
scaling is an important measure of how useful a solution technique is
when larger and larger values of $P$ are used to solve the
problem. Ideally, the time to solution for the the problem will
decrease linearly as a function of $P$, permitting a faster time to
solution using the maximum available resources thus making sure all
$P$ are positively affecting the runtime of the problem.

For strong scaling, the objective is to then decrease the runtime of a
fixed size global problem as a linear function of $P$. We can then
define the strong scaling \textit{absolute efficiency} as:
\begin{equation}
  \eta_{strong}(N,P) = \frac{1}{P} \frac{T(N,1)}{T(N,P)}\:,
  \label{eq:strong_scaling_absolute}
\end{equation}
where $T(N,P)$ is the wall time for a computation of global size $N$
using $P$ processes. Note here that this particular definition is
measured with respect to a serial computation. On leadership class
machines, it is often the case that a problem size large enough to
effectively execute on $O(10,000)$ cores or larger is significantly
larger than any problem that may be solved by a single core due to
memory restrictions. Therefore, we consider absolute scaling for a
base case where $Q$ processes are used instead of a serial
computation:
\begin{equation}
  \eta_{strong}(N,P|Q) = \frac{Q}{P} \frac{T(N,Q)}{T(N,P)}\:.
  \label{eq:strong_scaling_absolute_ref}
\end{equation}
The absolute efficiency measure presented here only considers the
total time to solution. Often, that total time to solution may be
changing for an iterative method due to the fact that the number of
iterations required to converge may be changing as a function of
$P$. This will often happen in cases where modifying $N$ changes the
spectral character of the problem, often making it stiffer and
therefore more difficult to solve when $N$ increases. Therefore we
also consider \textit{algorithmic efficiency} from Keyes' work which
considers this potential change in iteration count:
\begin{equation}
  \eta_{alg}(N,P|Q) = \frac{I(N,Q)}{I(N,P)}\:,
  \label{eq:algorithmic_efficiency}
\end{equation}
where $I(N,P)$ is the number of iterations required to converge a
problem of global size $N$ computed using $P$ processes. We can then
adjust the absolute efficiency using this new algorithmic efficiency
in order to compute the parallel scalability of a single iteration
such that algorithmic effects may be neglected. We define the
\textit{implementation efficiency} as:
\begin{equation}
  \eta_{impl}(N,P|Q) = \frac{\eta}{\eta_{alg}}\:,
  \label{eq:implementation efficiency}
\end{equation}
using the absolute efficiency in the numerator, giving an effective
measure of scalability of a single iteration of the algorithm.

@@

%%---------------------------------------------------------------------------%%
\section{Leadership-Class Parallel Scaling Studies\ }
\label{sec:leadership_scaling_studies}
Using the parallel performance metrics presented in the previous
section, we will next performance various scaling studies using an
implementation of the parallel MCSA algorithm to assess the quality of
the algorithm. For these scaling studies, large problems at high
levels of concurrency will be solved at a leadership-class computing
facility. In this section, we will outline the parallel machine on
which the calculations were performed and then perform a series a
studies that identify both strengths and weaknesses of the parallel
algorithm. Both pure domain decomposition and MSOD parallel
performance will be explored and compared to performance results for
the same problems using a set of production parallel Krylov solvers.

\subsection{Titan }
\label{subsec:titan}
All parallel scaling studies presented in this work were performed on
the Titan Cray XK7 machine at the Oak Ridge Leadership Computing
Facility located at Oak Ridge National
Laboratory\footnote{Specifications presented for Titan in the section
  were gathered from the Oak Ridge Leadership Computing Facility at
  \url{http://www.olcf.ornl.gov/computing-resources/titan-cray-xk7/}}. Titan
is a heterogeneous architecture machine combining both CPU and GPU
components to achieve a theoretical peak performance of more than 20
petaflops. The technical specifications for Titan are given in
Table~\ref{tab:titan_hardware} and spefically for only the CPU portion
of the machine.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      Processor & 16 core AMD \\
      Nodes & 18,688 AMD Opterons \\
      Cores/Node & 16 \\
      Total Cores & 299,008 \\
      Memory/Node & 32 GB \\
      Memory/Core & 2 GB \\
      Interconnect & Gemini \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Titan Cray XK7 hardware specifications.}}
  \label{tab:titan_hardware}
\end{table}

For the scaling studies presented in this work, only the CPU
components specified in of Titan were utilized. This is due to the
fact that the implementations of the parallel MCSA algorithm utilized
code written only with the Message Passing Interface (MPI) and
therefore could not take advantage of the GPU accelerators. Even
considering just the CPU portion of titan, nearly 300,000 CPUs are
available to test the parallel performance of the algorithm at high
levels of concurrency.

@@

In our description of Brunner and Brantley's algorithm as applied to
the Neumann-Ulam method, we noted two free parameters that are
user-defined at runtime. The first is the static data buffer size for
communicating histories between neighboring domains and the second is
the frequency by which a domain checks for incoming information from
neighboring domain, both of which may potentially affect latency and
load balancing. Changing the static history buffer size parameter
affects the scalability of the algorithm in two ways. If the buffer
size is increased, the buffer takes longer to fill and is therefore
sent less frequently, potentially causing scalability bottlenecks
where a neighboring domain is idle and waiting for more work to do but
is waiting for another domain to complete work. In addition, if the
buffer size is too large, the network may become saturated with these
very large data packets, slowing down performance. If the buffer size
is too small, too many history messages are sent, not providing enough
data for a neighboring process to work on to make up for the time
spent communicating. Considering the second parameter, when the check
frequency period is increased, more histories are processed locally
before looking to do work elsewhere. If this value becomes too large,
incoming data buffers may not be cleared frequently enough and cause a
bottle neck in the domain-to-domain communications. If this value is
too small, too much time is spent looking for work instead of actually
performing it.

To assess how these parameters affect the performance of the parallel
Monte Carlo algorithm on the Titan machine, a parameter study was
performed using the neutron diffusion problem on 64 cores (4
nodes). Each parameter was varied between 16 and 4096 and the time to
solve the problem recorded. The results of this study are presented in
Figure~\ref{fig:titan_comm_parameters}. Compared to the results of
Brunner and Brantley for the same type of study, we note qualitatively
similar results with the general observation that a larger history
buffer check period is better than a shorter one. The larger the check
period, the more on-process work is done before checking
messages. Second, there was not a significant variation observed in
the history buffer size parameter for a given check
frequency.

@@

\subsection{Pure Domain Decomposition}
\label{subsec:pure_domain_decomp}

\subsubsection{Strong Scaling}
\label{subsubsec:pure_strong}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_pure_strong.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition strong scaling absolute
      efficiency.}  \textit{Super-linear speed-up from memory
      thrashing in base case. MCLS is an order of magnitude slower
      arithmetically, causing the higher efficiency.}}
  \label{fig:titan_pure_strong}
\end{figure}

\subsubsection{Weak Scaling}
\label{subsubsec:pure_weak}

@@

@@

>>>>

%%---------------------------------------------------------------------------%%
\section{Parallel MCSA\ }
\label{sec:parallel_mcsa}
With the parallel adjoint Neumann-Ulam solver implementation described
aboive, the parallel implementation of an MCSA iteration is
trivial. Recall the MCSA iteration procedure presented again here for
clarity.
\begin{subequations}
  \begin{gather}
    \label{eq:mcsa_r1}
    \ve{x}^{k+1/2} = \ve{x}^k + \ve{r}^k\:,\\
    \label{eq:mcsa_r2}
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\
    \label{eq:mcsa_r3}
    \ve{A}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:,\\
    \label{eq:mcsa_r4}
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}\:,\\
    \label{eq:mcsa_r5}
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:,
  \end{gather}
\end{subequations}
In \S\ref{sec:parallel_krylov_methods} we discuss parallel matrix and
vector operations as utilized in conventional projection methods. We
utilize these here for the parallel MCSA implementation for the
computations required in Eqs~(\ref{eq:mcsa_r1}), (\ref{eq:mcsa_r2}),
(\ref{eq:mcsa_r4}), and (\ref{eq:mcsa_r5}). In the first step, a
parallel matrix-vector multiply is used to apply the split operator to
the previous iterate's solution. A parallel vector update is then
performed with the source vector to arrive at the initial iteration
guess. In the next step, the residual is computed by the same
operations where now the operator is applied to the solution guess
with a parallel matrix-vector multiply and then a parallel vector
update with the source vector is performed. Once the correction is
computed with a parallel adjoint Neumann-Ulam solve, this correction
is applied to the guess with a parallel vector update to get the new
iteration solution. Additionally, as given by
Eq~(\ref{eq:mcsa_stopping_criteria}), 2 parallel vector reductions
will be required to check the stopping criteria: one initially to
compute the infinity norm of the source vector, and another at every
iteration to compute the infinity norm of the residual vector.

To parallelize the solution of Eq~(\ref{eq:mcsa_r3}), we apply the
MSOD Neumann-Ulam algorithm presented in the previous section. The
problem is potentially replicated with overlap and the Monte Carlo
procedure returns a correction vector, $\delta \ve{x}$, with the same
parallel decomposition as the input linear problem. As was the case
for generating the replications required for a multiple set
implementation, the explict matrix and vector operations required for
the rest of an MCSA iteration occur only on the subset of processors
containing the primary set. Only the Monte Carlo data structures exist
on the rest of the processors in the problem as replication of the
other operations does not generate any more information that could be
used to accelerate the time to solution.

@@

Of primary interest to the nuclear engineer is the \textit{strong
  scaling} performance of an algorithm
\citep{siegel_analysis_2012}. In this type of scaling study, the
global problem size is fixed and the local problem size changes with
the number of processes in the calculation. In this case, as $P$
increases, $N$ is fixed, thus reducing the local problem size and
performance by having a larger ratio of communication to computational
work. This type of scaling is important because typically the global
problem size is fixed due to the problem of interest. Consider the
desire for a high fidelity neutronics simulation of the entire core of
a nuclear reactor. In this case, the global problem is known from the
geometry definition including the number of fuel assemblies and the
geometry of those assemblies all the way down to the resolution of the
space, angle, and energy discretizations required in order to capture
physical phenomena of interest. Because the global problem size is
known a priori, strong scaling is an important measure of how useful a
solution technique is when larger and larger values of $P$ are used to
solve the problem. Ideally, the time to solution for the the problem
will decrease linearly as a function of $P$, permitting a faster time
to solution using the maximum available resources and thus making sure
all $P$ are positively affecting the runtime of the problem.

For strong scaling, the objective is to then decrease the runtime of a
fixed size global problem as a linear function of $P$. Based on this,
we can define the strong scaling \textit{absolute efficiency} as:
\begin{equation}
  \eta_{strong}(N,P) = \frac{1}{P} \frac{T(N,1)}{T(N,P)}\:,
  \label{eq:strong_scaling_absolute}
\end{equation}
where $T(N,P)$ is the wall time for a computation of global size $N$
using $P$ processes. Note here that this particular definition is
measured with respect to a serial computation. On leadership class
machines, it is often the case that a problem size large enough to
effectively execute on $O(10,000)$ cores or larger is significantly
larger than any problem that may be solved by a single core due to
memory restrictions. Therefore, we consider absolute scaling for a
base case where $Q$ processes are used instead of a serial
computation:
\begin{equation}
  \eta_{strong}(N,P|Q) = \frac{Q}{P} \frac{T(N,Q)}{T(N,P)}\:.
  \label{eq:strong_scaling_absolute_ref}
\end{equation}
The absolute efficiency measure presented here only considers the
total time to solution. Often, that total time to solution may be
changing for an iterative method due to the fact that the number of
iterations required to converge may be changing as a function of
$P$. This will often happen in cases where modifying $N$ changes the
spectral character of the problem, often making it stiffer and
therefore more difficult to solve when $N$ increases. Therefore we
also consider \textit{algorithmic efficiency} from Keyes' work which
considers this potential change in iteration count:
\begin{equation}
  \eta_{alg}(N,P|Q) = \frac{I(N,Q)}{I(N,P)}\:,
  \label{eq:algorithmic_efficiency}
\end{equation}
where $I(N,P)$ is the number of iterations required to converge a
problem of global size $N$ computed using $P$ processes. We can then
adjust the absolute efficiency using this new algorithmic efficiency
in order to compute the parallel scalability of a single iteration
such that algorithmic effects may be neglected. We define the
\textit{implementation efficiency} as:
\begin{equation}
  \eta_{impl}(N,P|Q) = \frac{\eta}{\eta_{alg}}\:,
  \label{eq:implementation efficiency}
\end{equation}
using the absolute efficiency in the numerator, giving an effective
measure of scalability of a single iteration of the algorithm.

@@

%%---------------------------------------------------------------------------%%
\section{Leadership-Class Parallel Scaling Studies\ }
\label{sec:leadership_scaling_studies}
Using the parallel performance metrics presented in the previous
section, we will next performance various scaling studies using an
implementation of the parallel MCSA algorithm to assess the quality of
the algorithm using the neutron diffusion problem presented in
Appendix~\ref{chap:diffusion_problem}. For these scaling studies,
large problems at high levels of concurrency will be solved at a
leadership-class computing facility. In this section, we will outline
the parallel machine on which the calculations were performed and then
present a series of numerical studies that identify both strengths and
weaknesses of the parallel algorithm. Both pure domain decomposition
and MSOD parallel performance will be explored and compared to
performance results for the same problems using a set of production
parallel Krylov solvers. For the scaling results presented here, all
parallel Monte Carlo results were computed using the Monte Carlo
Linear Solvers\footnote{MCLS is an open source library and can be
  downloaded at \url{https://github.com/sslattery/MCLS}} (MCLS)
library generated specifically for this work while the conjugate
gradient and GMRES Krylov solvers used are part of the Belos linear
solver package within the Trilinos framework
\citep{heroux_overview_2005}. In general, the Belos package produces
solutions to the neutron diffusion problem over an order of magnitude
faster than the MCLS library for serial computations. Therefore, when
we directly compare these two codes, the MCLS parallel efficiencies
will appear artificially inflated relative to the Belos efficiencies
due to the fact that the MCLS ratio of work to communication will be
much larger than that for the Belos calculations.

\subsection{Titan }
\label{subsec:titan}
All parallel scaling studies presented in this work were performed on
the Titan Cray XK7 machine at the Oak Ridge Leadership Computing
Facility located at Oak Ridge National Laboratory. Titan is a
heterogeneous architecture machine combining both CPU and GPU
components to achieve a theoretical peak performance of more than 20
petaflops. The technical specifications for Titan are given in
Table~\ref{tab:titan_hardware} and specifically for only the CPU
portion of the machine.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      Processor & 16 core AMD \\
      Nodes & 18,688 AMD Opterons \\
      Cores/Node & 16 \\
      Total Cores & 299,008 \\
      Memory/Node & 32 GB \\
      Memory/Core & 2 GB \\
      Interconnect & Gemini \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Titan Cray XK7 hardware specifications.}
    \textit{Specifications presented for Titan in this table were
      gathered from the Oak Ridge Leadership Computing Facility at}
    \url{http://www.olcf.ornl.gov/computing-resources/titan-cray-xk7/}}
  \label{tab:titan_hardware}
\end{table}

For the scaling studies presented in this work, only the CPU
components of Titan specified in Table~\ref{tab:titan_hardware} were
utilized. This is due to the fact that the implementations of the
parallel MCSA algorithm utilized code written only with the Message
Passing Interface (MPI) and therefore could not take advantage of the
GPU accelerators. Even considering just the CPU portion of Titan,
nearly 300,000 CPUs are available to test the parallel performance of
the MCSA algorithm at high levels of concurrency.

@@

In our description of Brunner and Brantley's algorithm as applied to
the Neumann-Ulam method, we noted two free parameters that are
user-defined at runtime. The first is the static data buffer size for
communicating histories between neighboring domains and the second is
the frequency by which a domain checks for incoming information from
neighboring domain, both of which may potentially affect latency and
load balancing. Changing the static history buffer size parameter
affects the scalability of the algorithm in two ways. If the buffer
size is increased, the buffer takes longer to fill and is therefore
sent less frequently, potentially causing load balancing bottlenecks
where a neighboring domain is idle and waiting for more work to do but
is waiting for another domain to complete work. In addition, if the
buffer size is too large, the network may become saturated with these
very large data packets, slowing down performance. If the buffer size
is too small, too many history messages are sent, not providing enough
data for a neighboring process to work on to make up for the time
spent communicating. Considering the second parameter, when the
message check frequency period is increased, more histories are
processed locally before looking to do work elsewhere. If this value
becomes too large, incoming data buffers may not be cleared frequently
enough and cause decreased performance in the domain-to-domain
communications. If this value is too small, too much time is spent
looking for work instead of actually performing it.

To assess how these parameters affect the performance of the parallel
Monte Carlo algorithm on the Titan machine, a parameter study was
performed using the neutron diffusion problem on 64 cores (4
nodes). This problem had a reasonably well conditioned spectral radius
of 0.78 with a global problem size of \sn{1.6}{7} DOFs with one
stochastic history used per DOF in the MCSA Neumann-Ulam used to
compute the correction for the acceleration. Each parameter was varied
between 16 and 4096 and the time to solve the problem recorded. The
results of this study are presented in
Figure~\ref{fig:titan_comm_parameters}. Compared to the results of
Brunner and Brantley for the same type of study, we note qualitatively
similar results with the general observation that a larger history
buffer check period is better than a shorter one. The larger the check
period, the more on-process work is done before checking
messages. Second, there was not a significant variation observed in
the history buffer size parameter for a given check frequency.

@@

\subsection{Pure Domain Decomposition Scaling}
\label{subsec:pure_domain_decomp}

For the first set of scaling studies, we consider the parallel MCSA
algorithm where MSOD has not been used in the parallel Neumann-Ulam
algorithm. We refer to this case as pure domain decomposition where no
overlap or replication has been leveraged. Such a test will isolate
specifically the performance of Brunner and Brantley's algorithm as
applied to the Neumann-Ulam algorithm. For the strong scaling study,
the global problem size was fixed at \sn{1.6}{7} DOFs for all solvers
and the 16 core run used as the base case. MCSA was used again with
one stochastic history used per DOF in the Neumann-Ulam solve to
compute the correction for the
acceleration. Figure~\ref{fig:titan_pure_strong} gives the results of
the this study with the given absolute efficiencies computed using
Eq~(\ref{eq:strong_scaling_absolute}).

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_pure_strong.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition strong scaling absolute
      efficiency.}  \textit{Super-linear speed-up is from memory
      thrashing in the base case. MCLS is an order of magnitude slower
      arithmetically, causing the higher efficiency.}}
  \label{fig:titan_pure_strong}
\end{figure}

There are several key features of Figure~\ref{fig:titan_pure_strong}
worth noting. First, up to around 5,000 cores, all methods experience
a super-linear improvement in parallel efficiencies. For all methods,
this increase is not a function of the algorithmic efficiency as those
computed using Eq~(\ref{eq:algorithmic_efficiency}) remained at unity,
meaning that all computations converged in the same number of
iterations relative to the base case\footnote{When the algorithmic
  efficiency remains at unity for all calculations, the implementation
  efficiency is equivalent to the absolute efficiency and therefore is
  not presented here.}. Instead, what is happening here and was
observed by Brunner and Brantley in their work is that the large
problem used to scale to O(10,000) cores was large enough that memory
thrashing was occuring in the 16 core base case. As the number of
cores in the strong scaling study was increased and the local problem
size decreased, more of the local problem fit into the cache of the
system, thus causing improved serial performance and therefore
enhanced runtimes. In this region of super-linear efficiencies, we
also observe larger efficiencies for GMRES as compared to conjugate
gradient. This occurs due to the fact that conjugate gradient has a
static set of operations performed at every iteration while GMRES
operates on a continually growing subspace. In addition, the
orthogonalization operations required to generate that subspace
consume much more time than the basic conjugate gradient operations,
causing a larger serial runtime and subsequently higher
efficiencies. For the same reasons, we also observe higher
efficiencies for the MCSA implementation due to the serial performance
of the implementation.

Beyond around 5,000 cores, the global problem size selected for this
strong scaling study causes the local problem size to shrink enough
that parallel communication costs begin to overtake arithmetic costs
on process. This takeover causes this effective strong scaling wall
that is observed in all methods. Using any more cores to solve this
particular global problem is not an efficient use of those parallel
resources. Furthermore, the fact that we see a very similar trends in
efficiencies for all methods means that they are bound by the same
parallel performance limitations in a strong scaling environment. This
is important as we can expect the same qualitative scaling performance
from MCSA onc arithmetic optimization has been performed to improve to
serial performance of the implementation. In addition, these results
are valuable in that the show MCSA, when properly parallelized, has
the ability to be competetive on leadership-class computing platforms
with conventional subspace methods with good strong scaling observed
in the pure domain decomposition case to O(10,000) cores.

Next, we consider the weak scaling performance of the algorithm. In
this case, we fix the local problem size at \sn{4}{4} DOFs and
increase the number of processors used in the
simulation. Figure~\ref{fig:titan_weak_absolute} gives the weak
scaling absolute efficiencies for this problem as computed by
Eq~(\ref{eq:weak_scaling_absolute}). In general, the relative trends
when comparing MCSA to the Krylov methods are the same with MCSA and
GMRES scaling better than the conjugate gradient method due to the
increased amount of arithmetic work on-process. In addition, it is
important to note that the weak scaling of MCSA when properly
parallelized remains competetive with the Krylov methods keeping in
mind the artificial inflation of efficiencies.

@@

As an additional means of comparison between MCSA and the Krylov
methods, we compute the algorithmic efficiencies of the calculations
using Eq~(\ref{eq:algorithmic_efficiency}). As shown in
Figure~\ref{fig:titan_weak_algorithmic}, the algorithm efficiency of
the Krylov algorithm improves as a function of the number of cores in
the problem due to the fact that the number of iterations required to
converge decreased from 20 in the base case to 19 and then to 18 for
the final set of computations. When this reduction in iterations is
used to account for the observed runtimes and computed absolute
efficiencies for weak scaling we end up with the implementation
efficiencies computed using Eq~(\ref{eq:implementation_efficiency})
given in Figure~\ref{fig:titan_weak_implementation}. Because MCSA
maintains the same iterative performance as a function of global
problem size in the weak scaling study, the implementation
efficiencies compare more favorably to the Krylov methods. In effect,
we are measuring the parallel performance of a single iteration for
each method by looking at these implementation efficieinces. As the
global problem size increases, MCSA is less prone to a drop in
parallel efficiency. When optimization has been completed and runtimes
potentially approach those of the Krylov methods, this comparison may
not be as favorable. Overall, weak scaling performance the parallel
MCSA algorithm is excellent for the pure domain decomposition case
with absolute parallel efficiencies of over 80\% observed at over
32,000 cores.

@@

In addition to comparing MCSA scaling to conventional methods, we also
motivate avoiding any type of master/slave algorithms at these high
levels of concurrency by comparing the binary communication tree
implementation used in the previous results with a parallel MCSA
implementation that uses the master/slave scheme to determine the end
of a history transport stage as shown in
Figure~\ref{fig:master_comm_tree}
instead. Figure~\ref{fig:titan_strong_bvsm} provides this comparison
using the strong scaling study while Figure~\ref{fig:titan_weak_bvsm}
gives the results of the comparison for the weak scaling study. In
both cases, the deficiencies caused by using the master/slave scheme
for transport completion are obvious. At around 1,000 cores, both the
strong and weak scaling performance seriously degrades while the
binary tree pattern maintains execellent performance. At this number
of cores, any extra work given to the master process, even if it
amounts to summing a handful of integers and checking for a few extra
messages as in this algorithm, this extra work grows enough to create
serious load imbalance and reduce performance.
